---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
rawData <- read.csv("activity.csv")
dataSample <- head(rawData)
print(str(rawData))
print(dataSample, type="html")
```


## What is mean total number of steps taken per day?

```{r}
library(dplyr)
dailySteps <- rawData %>% group_by(date) %>% 
    summarise(total_steps = sum(steps, na.rm=TRUE))
hist(dailySteps$total_steps, main="Number of Steps per Day", xlab = "Step Count",
     ylab="Number of Days")
```

```{r}
meanStepsPerDay <- mean(dailySteps$total_steps)
medianStepsPerDay <- median(dailySteps$total_steps)
```

The mean total number of steps per data is `r meanStepsPerDay`.  
The median total number of steps per day is `r medianStepsPerDay`.  

## What is the average daily activity pattern?

```{r}
intervalSteps <- rawData %>% group_by(interval) %>% 
    summarise(average_steps = mean(steps, na.rm=TRUE))
maxStepsInterval <- intervalSteps$interval[which.max(intervalSteps$average_steps)]
maxSteps <- round(max(intervalSteps$average_steps),0)
with(intervalSteps, plot(interval, average_steps, type="l",
                         main="Daily Activity Pattern", 
                         xlab="Time Interval (mins)",
                         ylab="Average Steps"))
```

The `r maxStepsInterval`th interval contains the most steps, `r maxSteps`.  

## Imputing missing values

```{r}
missingDataCount <- sapply(rawData, function(x) sum(is.na(x)))
print(missingDataCount, type="html")
```

_Replace missing data with mean steps for the interval averaged accross the days_
```{r}
imputedData <- rawData
for (i in 1:dim(imputedData)[1]) {
    if (is.na(imputedData[i,"steps"])) {
        imputedData[i,"steps"] <- intervalSteps[intervalSteps$interval == imputedData[i, "interval"],
                                                "average_steps"]
    }
}
```

```{r}
imputedDailySteps <- imputedData %>% group_by(date) %>% 
    summarise(total_steps = sum(steps, na.rm=TRUE))
hist(imputedDailySteps$total_steps, main="Number of Steps per Day", xlab = "Step Count",
     ylab="Number of Days")
```

```{r}
imputedMeanStepsPerDay <- mean(imputedDailySteps$total_steps)
rawMeanDiff <- round(imputedMeanStepsPerDay - meanStepsPerDay,0)
imputedMedianStepsPerDay <- median(imputedDailySteps$total_steps)
rawMedianDiff <- round(imputedMedianStepsPerDay - medianStepsPerDay,0)
```

The mean total number of steps per data is `r imputedMeanStepsPerDay`.
Difference from raw mean = `r rawMeanDiff`
The median total number of steps per day is `r imputedMedianStepsPerDay`. 
Difference from raw median = `r rawMedianDiff`

Due to imputing, the totals will increase as initially the na will be ignored.

## Are there differences in activity patterns between weekdays and weekends?
